---
title: "Seedlings recruitment analysis of Solarik et al. 2017 paper."
author: "Kevin Cazelles and Kevin Solarik"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: readable
    toc: yes
    toc_float: yes
vignette: >
  %\VignetteIndexEntry{Seedlings recruitment}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---



# Abstract


This vignette aims at guiding the reader through the different steps of
the analysis we carried out in Solarik et al. XXXX. Abstract is as follows:


## Data

Part of the datasets published [here]()[^tobeadded].


### Sites

```{r map, message=FALSE, echo=FALSE}
datS <- data.frame(
  id = c('abi', 'bic', 'sut'),
  name = c('Abitibi-Temiscaminque', 'Le Bic', 'Sutton'),
  lon = c(-79.401219, -68.817719, -72.541297),
  lat = c(48.162539, 48.333619, 45.112803),
  plots= c(NA, NA, NA)
)
##
rownames(datS) <- datS$id
spSite <- sp::SpatialPointsDataFrame(
    datS[,c('lon', 'lat')],
    data = datS,
    proj4string = sp::CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")
  )
mapview::mapview(spSite)
```



## Tree species of the study

| abbr1 | abbr2 | Scientific Names      | TSN    |abi  |bic  |sut  |
|:------|:------|:----------------------|:-------|:----|:----|:----|
| BF    | ABBA  | Abies balsamea        | 18032  |TRUE |TRUE |TRUE |
| RM    | ACRU  | Acer rubrum           | 28728  |TRUE |TRUE |TRUE |
| SM    | ACSA  | Acer saccharum        | 28731  |TRUE |TRUE |TRUE |
| YB    | BEAL  | Betula alleghaniensis | 19481  |TRUE |TRUE |TRUE |
| WB    | BEPA  | Betula papyrifera     | 19489  |TRUE |TRUE |TRUE |
| AB    | FAGR  | Fagus grandifolia     | 19462  |TRUE |TRUE |TRUE |
| AS    | POTR  | Populus tremuloides   | 195773 |TRUE |TRUE |TRUE |

: abbreviations, names, TSN (Taxonomic Serial Number, viti the [Integrated
  Taxonomic Information System website](https://www.itis.gov/pdf/faq_itis_tsn.pdf)
  for more details about TSN) of the tree species considered and their
  respective presence in the three sites of the study.




## Statistical model

### Model

### Parameters

| Ecological process  |abbr  |number of parameters|
|:--------------------|:-----|:-------------------|
| dispersion          |disp  |2                   |
| neighborhood effect |neigh |5                   |
| favorability        |favo  |1                   |



## Overall strategy of our analysis

Likelihood optimization using simulated annealing
(implemented by the [GenSA](https://cran.r-project.org/web/packages/GenSA/index.html) package).
For different models (40 models for each combination site x species x year x age).




[^tobeadded]: Ref To be added once the paper published











<!--
## File index


```{r, eval=FALSE}
fl_simu <- data/simu_all.Rds
```

```{r, eval=FALSE, echo=FALSE}
setwd("/Users/KevCaz/Dropbox/Kev_code")
```





# Formatting the data

Before starting, do not forget to chose properly your working space.
The code lines below creates three folder where data, results and figures will
be stored.

```{r, eval=FALSE}
graphics.off()
rm(list=ls())
sapply(c("data", "res", "fig"), dir.create)
```

## Downloading the data

Would be either url or included in the packages (either `.csv` or `rds` files).


## Installing the package

For convenience, all the functions are gathered in a R package, the `devtools` package makes the installation of github packages easy.

```{r, eval=FALSE}
if (!require("devtools")) install.packages("devtools")
devtools::install_github("KevCaz/seedlingsRecruitmentAnalysis")
```


### Formatting raw data

Columns description:

```{r, eval=FALSE}
colnames()
```

This section will be completed once the paper published.

```{r, eval=FALSE}
spe_trees_abi <- c("ABBA", "ACRU", "ACSA", "BEPA")
spe_trees_bic <- c("ABBA", "ACSA", "ACRU", "BEPA", "POTR")
spe_trees_sut <- c("ABBA", "ACSA", "BEAL", "BEPA", "FAGR")
```


### Abitbi (abi)

```{r, eval=FALSE}
getDataReady(
  fl_trees = "data/abitbitrees.csv",
  fl_regen = "data/abitibiregen.csv",
  fl_regen2 = "data/abitibiregen2016.csv",
  treesp = spe_trees_abi,
  abr_site = "abi",
  dist_buffer = 20
)
```

### Le Bic, Rinouksi (bic)

```{r, eval=FALSE}
getDataReady(
  fl_trees = "data/rimouskitrees.csv",
  fl_regen = "data/rimouskiregen.csv",
  fl_regen2 = "data/rimouskiregen2016.csv",
  treesp =  spe_trees_bic,
  abr_site = "bic",
  dist_buffer = 20
)
```


### Sutton (sut)

```{r, eval=FALSE}
getDataReady(
  fl_trees = "data/suttontrees.csv",
  fl_regen = "data/suttonregen.csv",
  fl_regen2 = "data/suttonregen2016.csv",
  treesp = spe_trees_sut,
  abr_site = "sut",
  dist_buffer = 20
)
```



## Generating the file describing all simulations


```{r, eval=FALSE}
ls_site <- list(spe_trees_abi, spe_trees_bic, spe_trees_sut)
szs <- unlist(lapply(ls_site, length))
##
grd <- expand.grid(
  disp = c(0,1,2,3,4),
  favo = c(TRUE, FALSE),
  neigh = c(TRUE, FALSE),
  pz = c(TRUE, FALSE),
  age = c("1", "2"),
  year = c(2015, 2016),
  tree = unlist(ls_site)
   ) %T>% str
##
simu <- cbind(site=rep(c("abi", "bic", "sut"), szs*2*80), grd[,c(7,6,5,4,1:3)])
saveRDS(simu, file = fl_simu)
```

To import the simulation file, use the following:

```{r, eval=FALSE}
simu <- readRDS(file="fl_simu")
```








# Parameters estimation

## Parameters ranges and estimation

```{r, eval=FALSE, echo=FALSE}
getParameters() %>% as.data.frame %>%
  knitr::kable(caption = 'Parameters values', booktabs = TRUE)
```

```{r, eval=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```


## Likelihood optimization

For all simulations, we apply the same analysis and proceed as follows:

```{r, eval=FALSE}
for (i in 1:nrowm(simu)){
  launchIt(i, simufile = "data/simu_all.Rds", mxt = 1000,
    record = paste0("test", adjustString(i,4), ".txt"), quiet=FALSE)
}
```

We benefited from calculus facilities (Mammoth super computer, Compute Canada).



## Convergence

To check which models converged:

```{r, eval=FALSE}
val <- as.numeric(gsub(list.files("res"), pat = "\\D", rep =""))
rerun <- which(!(1:1120 %in% val)) %T>% saveRDS("rerun.Rds")
```

```{r, eval=FALSE}
script_ci()
```


## Results of the GenSA analyses

```{r, eval=FALSE}
pathestim <- "~/Pending/kevSol/"
tmp <- read.csv(paste0(pathestim, "estim.csv"))
```

## Mean dispersal distance

```{r, eval=FALSE}
tmp$mean_disp_dist <- NA
```

```{r, eval=FALSE}
for (i in 1:nrow(tmp)){
  if (tmp$disp[i] & !is.na(tmp$shap[i]) & !is.na(tmp$scal[i])){
    if (tmp$disp[i]>2){
      nm <- "exponential_power"
    } else {
      nm <- "lognormal"
    }
    tmp$mean_disp_dist[i] <- meanDispDist(nm, tmp$scal[i], shap = tmp$shap[i])
  }
}
```


## McFadden's pseudo R-squared

```{r, eval=FALSE}
nrep <- nrow(unique(tmp[,c("site","tree","age")]))
##
tmp$McFad_adj$aic <- 2*tmp$nbpar - 2*tmp$lik
tmp$McFad_adj <- NA
tmp$best <- FALSE
##
nmod = 40
idrm <- c(
  subset(tmp, site=="abi" & tree=="ACRU" & year==2016 & age==1)$X,
  subset(tmp, site=="abi" & tree=="BEPA" & year==2016 & age==1)$X,
  subset(tmp, site=="abi" & tree=="BEPA" & year==2016 & age==2)$X,
  subset(tmp, site=="bic" & tree=="POTR" & year==2016 & age==1)$X,
  subset(tmp, site=="sut" & tree=="ABBA" & year==2015 & age==1)$X,
  subset(tmp, site=="sut" & tree=="ABBA" & year==2016 & age==1)$X
)
tmp <- tmp[-idrm,]
```


```{r, eval=FALSE}
# ti be removed
# | site |tree  | year | age |
# |:-----|:-----|:-----|:----|
# | abi  |ACCRU | 2016 | 1   |
# | abi  | BEPA | 2016 | 1   |
# | abi  | BEPA | 2016 | 2   |
# | bic  | POTR | 2016 | 1   |
# | sut  | ABBA | 2015 | 1   |
# | sut  | ABBA | 2016 | 1   |
```



```{r, eval=FALSE}
for (i in 1:(nrow(tmp)/nmod)){
  lag <- (i-1)*nmod
  rg <- lag + 1:nmod
  id <- lag + which(tmp$nbpar[rg]==1)
  print(tmp$lik[id])
  if (tmp$lik[id]>-0.001) print(tmp[id,1:5])
  tmp$McFad_adj[rg] <- 1 - (tmp$lik[rg]-(tmp$nbpar[rg]-1))/(tmp$lik[id])
  tmp$best[lag+which.max(tmp$McFad_adj[rg])] <- TRUE
}
##
tmp[tmp$best, c("McFad_adj")]*100
tmp[tmp$best, c("mean_disp_dist")]
tmp[tmp$best,] %>% kable(digits=4) %>% cat(sep="\n", file="~/Pending/kevSol/best_mod.md")
## produce table
saveRDS(tmp[tmp$best,], file="~/Pending/kevSol/resbest_all.Rds")
write.csv(tmp, file="~/Pending/kevSol/res_all.csv")
write.csv(tmp[tmp$best,], file="~/Pending/kevSol/resbest_all.csv")
## Symbol table / per year
get_symb_mod(tmp, year=2015) %>% kable %>% cat(sep="\n", file="~/Pending/kevSol/table_all2015.md")
get_symb_mod(tmp, year=2016) %>% kable %>% cat(sep="\n", file="~/Pending/kevSol/table_all2016.md")


##-- 1-3 clip
tmpc <- tmp[which(tmp$disp%in%c(0,1,3)),]
tmpc$McFad_adj <- NA
tmpc$best <- FALSE
nmod <- 24
for (i in 1:(nrow(tmpc)/nmod)){
  lag <- (i-1)*nmod
  rg <- lag + 1:nmod
  id <- lag + which(tmpc$nbpar[rg]==1)
  tmpc$McFad_adj[rg] <- 1 - (tmpc$lik[rg]-(tmpc$nbpar[rg]-1))/(tmpc$lik[id])
  tmpc$best[lag+which.max(tmpc$McFad_adj[rg])] <- TRUE
}
##
get_symb_mod(tmpc, year=2015) %>% kable %>% cat(sep="\n", file="~/Pending/kevSol/table_clip2015.md")
get_symb_mod(tmpc, year=2016) %>% kable %>% cat(sep="\n", file="~/Pending/kevSol/table_clip2016.md")

saveRDS(tmpc[tmpc$best,], file="~/Pending/kevSol/resbest_clip.Rds")
write.csv(tmpc[tmpc$best,], file="~/Pending/kevSol/resbest_clip.csv")



##-- 2-4 non-clip
tmpc <- tmp[which(tmp$disp%in%c(0,2,4)),]
tmpc$McFad_adj <- NA
tmpc$best <- FALSE
for (i in 1:(nrow(tmpc)/nmod)){
  lag <- (i-1)*nmod
  rg <- lag + 1:nmod
  id <- lag + which(tmpc$nbpar[rg]==1)
  tmpc$McFad_adj[rg] <- 1 - (tmpc$lik[rg]-(tmpc$nbpar[rg]-1))/(tmpc$lik[id])
  tmpc$best[lag+which.max(tmpc$McFad_adj[rg])] <- TRUE
}
get_symb_mod(tmpc, year=2015) %>% kable %>% cat(sep="\n", file="~/Pending/kevSol/table_nonclip2015.md")
get_symb_mod(tmpc, year=2016) %>% kable %>% cat(sep="\n", file="~/Pending/kevSol/table_nonclip2016.md")
saveRDS(tmpc[tmpc$best,], file="~/Pending/kevSol/resbest_nonclip.Rds")
write.csv(tmpc[tmpc$best,], file="~/Pending/kevSol/resbest_nonclip.csv")
```

```{r, eval=FALSE}
##-- Converting the tables from md. to pdf and docx.
for (inp in list.files("~/Pending/kevSol/", pattern=".md$", full.names=T)){
  out <- gsub(inp, pat=".md", rep="")
  system(paste0("pandoc ", inp, " -o ", out, ".pdf"))
}
```


## Figures

### Kernels

```{r, eval=FALSE}
# use the best model to subset the tables
tall <- readRDS("~/Pending/kevSol/resbest_all.Rds")
figKernels(tall[tall$year==2015,], filename="~/Pending/kevSol/figall2015.pdf")
figKernels(tall[tall$year==2016,], filename="~/Pending/kevSol/figall2016.pdf")
``` -->
